D.I.A.G.R.A.M.: Digital Image Analysis for Graph Recognition And Modeling

### 2025-07-09 doc/article/article.tex
Push

@@

\subsubsection{Text Association}
It can be assumed that every piece of text is linked to one and only one element or relation. So, every bounding box containing text is linked to the nearest element - it being either a relation or an element. The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.\\

\subsubsection{Element-Text Association}
After that a text box has been assigned to an element, there's the need to understand if the text is either inside or outside the element and whether the text is actually referring to the given element or not. In order to do this, the distance between the two bounding boxes is computed - as the distance between their central point. If the distance is over a certain threshold, the association is kept; otherwise, it's discarded.\\

\subsubsection{Relation-Text Association}
After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself. The relation's bounding box is then split into three different sections. The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned to the text under consideration. \\

@@

\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on the first three datasets: flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets. This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types, which were the core of our task.

@@

Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.

@@

However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow.

@@

>>>>

\subsubsection{Arrow Retrieval}
Sometimes, the number of arrow's bodies, heads and tails don't match up. In particular, the following, trivial, equivalance doesn't hold true:
$$
	2\times\text{arrow bodies} = \text{arrow heads} + \text{arrow tails}
$$

The case in which, for a specific arrow, its head or its tail has not been found. In order to be able to be as error-robust as possible, we give the arrow - and a crop with more area around it - to the extraction net once again. If it's able to find both an head and a tail this time, we assume those to be the one's linked to the arrow. If more than one head or tail is found, the highest confidence result is kept - unless the confidence is lower than a certain threshold.

The other case is the one in which we have some heads and tails without any body. For each pair of head and tail, the area of the image linking them is cropped and given to the detection net. For every arrow's bounding box found - if any - we compute the overlap with both the head and tail's bounding boxes. If the overlap is lower than a certain threshold, the arrow's bounding box is discarded. If it's over it and if it's over the current arrow's confidence value - if present, in the case in which we found multiple valid candidates - it's momentarily assumed as the correct arrow. If no arrow fulfills any of the above requisite, we discard the couple as candidates. \\

	% Row 1
	\begin{subfigure}[b]{0.45\linewidth}
		\includegraphics[width=\linewidth]{retrieval1.jpg}
		\caption{A graph diagram. The upper rightmost self-arrow has been detected - but not its tail and head.}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\linewidth}
		\includegraphics[width=\linewidth]{retrieval2.png}
		\caption{The detection net's output, able to recognize the arrow's head and tail after passing only a crop of the region.}
	\end{subfigure}

@@

\subsubsection{Text Association}
It can be assumed that every piece of text is linked to one and only one element or relation. So, every bounding box containing text is linked to the nearest element - it being either a relation or a node. The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.\\

\subsubsection{Node-Text Association}
After that a text box has been assigned to a node, there's the need to understand if the text is either inside or outside of the target node and whether the text is actually referring to the given element or not. In order to do this, the overlap between the two bounding boxes is computed; the value is compared to an heuristically determined threshold. If it's higher, the association is kept and the text is determined to be inside. If it's lower, the distance between the two bounding boxes is computed - as the distance between their central point. If the distance is below a certain threshold, the association is kept and the text is regarded as outside the node; otherwise, it's discarded.\\

\subsubsection{Arrow-Text Association}
After a text box is assigned to an arrow, some computation is needed to fully understand how the text is associated to the arrow itself. The arrow is approximated by a line adjoining its head and its tail. The distance between the line and the text's bounding box is computed and compared with a threshold - if it's over it, the association is discarded. Otherwise, the line approximation is then split into three different sections - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two. The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration. \\

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{arrowsplit.jpg}
	\caption{A three-split of an arrow.}
	\label{fig:arrowsplit}
\end{figure}

@@

\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on the first three datasets: flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets. This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types, which were the core of our task. \\

@@

Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}. \\

@@

However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow (see Figure~\ref{fig:selfarrowdc}).

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=150pt]{selfarrowdc.jpg}
	\caption{An instance which explains why the self-arrow's tail and head are not correctly recognized.}
	\label{fig:selfarrowdc}
\end{figure}
